---
title: "Project"
author: "Manideep"
date: "12/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(dlookr)
library(ggplot2)
library(rworldmap)
library(ggmap)
library(tidyverse)
library(car)
```

```{r}
dataset <- read.csv("data1.csv", header = TRUE)
dataset
```

````{r}
summary(dataset)
````


```{r}
str(dataset)
```


```{r}
nrow(dataset[!complete.cases(dataset),])
sum(is.na(dataset))
colSums(is.na(dataset))
```
```{r}
#remove rows with null values in 9 columns
dataset <- dataset[!is.na(dataset$Region),]
#impute missing diabetes value with 0
dataset[is.na(dataset)] <- 0

sum(is.na(dataset))
summary(dataset$Prevalence_raised_blood_pressure)
```
```{r}
dataset
```

```{r}
#log transformation
dataset$LogObesity <- log(dataset$Prevalence_obesity_adults+3)
dataset$LogMorbidObesity <- log(dataset$Prevalence_morbid_obesity_adults+3)
dataset$LogUnderwight <- log(dataset$Prevalence_underweight_adults+3)
dataset$LogWesternDiet <- log(dataset$Western_diet_score+3)
```

```{r}
#univariate analysis
ggplot(dataset, aes(x = Mean_BMI_adults, fill = Sex)) +geom_histogram(position = "dodge")
ggplot(dataset, aes(x = Prevalence_obesity_adults, color = "overweight")) +geom_density() + 
  geom_density(aes(x = Prevalence_underweight_adults, color = "underweight"))+
  scale_color_manual(name = "Weights", values = c("overweight" = "black","underweight" = "red"))
```

```{r fig.width==10, fig.height=8}
library(corrplot)
num_col <- dataset[, sapply(dataset, is.numeric)]
values_cor<-cor(num_col)
round(values_cor,2)
corrplot(cor(num_col, use="pairwise.complete.obs"), method = "circle")
```

```{r}
#multiple variables
dataset %>% filter(Superregion != 0) %>%
  ggplot(aes(x = Year, y = Diabetes_prevalence, color = Sex)) +geom_smooth() +facet_wrap(~Superregion)
ggplot(dataset, aes(x = Year, y = Prevalence_obesity_adults, color = Sex)) +geom_smooth() +facet_wrap(~Superregion)
```

```{r}
#multiple regression model for diabetes
formula <- Diabetes_prevalence ~ Sex+Year + Mean_BMI_adults +LogObesity +LogWesternDiet+LogUnderwight+LogMorbidObesity +
  Systolic_blood_pressure +  Prevalence_raised_blood_pressure  + Region + Superregion +                 
  Years_of_education  + Urbanisation   + Country                         
model <- lm(formula, data = dataset)
#model Summary
summary(model)
```

```{r}
# Plot for model 1
#plot(final_dataset_data.model1)
par(mfrow=c(2,2))
plot(model)
par(mfrow=c(1,1))
```
```{r}
#remove insignificant variables 

final_formula <- Diabetes_prevalence  ~ Sex + Year + LogObesity + 
  Prevalence_raised_blood_pressure  + Country

model1 <- lm(final_formula, data = dataset)
saveRDS(model1, file = "./model_1.rda")

summary(model1)
```
```{r}
new_val_dia=tibble(Sex=c('Female','Male') , Year=c(1981),
                   LogObesity=c(0.6),Prevalence_raised_blood_pressure=c(0.13),Country=c('Bahamas'))
diabetic_predict = predict.lm(model1, new_val_dia)
diabetic_predict
```
```{r}
# Calculating 95% of standardized residuals between -1.96 and 1.96 for outliers for model 2
dataset$fitted <- model1$fitted
dataset$residuals <- model1$residuals
dataset$standardized.residuals <- rstandard(model1)
possible.outliers_model_1 <- subset(dataset, standardized.residuals < -1.96 | standardized.residuals > 1.96)
possible.outliers_model_1
```
```{r}
plot(model1)
```
```{r}
# Cooks distance for influential cases for model 2
dataset$cooks <- cooks.distance(model1)
plot(sort(dataset$cooks, decreasing=TRUE))
```

````{r}
max(dataset$cooks)
````
```{r}
#Annova tests
res_aov_country <- aov(Diabetes_prevalence ~ Country, data = dataset )
summary(res_aov_country)
res_aov_region <- aov(Diabetes_prevalence ~ Region, data = dataset)
```

````{r}
anova(model, model1)
```
```{r}
###plot the residuals
plot(dataset$Diabetes_prevalence, model1$residuals)
abline(h=0,lty=2)
hist(model1$residuals, main = "Histograms of residuals")
```

``````{r}
#regression for blood pressure
formula__bp <- Prevalence_raised_blood_pressure ~ Diabetes_prevalence + Sex + Year  +
  Systolic_blood_pressure   + Country


model_bp <- lm(formula_bp, data = dataset)
summary(model_bp)
````
```{r}
###plot the residuals
plot(dataset$Prevalence_raised_blood_pressure, model_bp$residuals)
abline(h=0,lty=2)
hist(model_bp$residuals, main = "Histograms of residuals")
```
````{r}
plot(model_bp)
````
```{r}
# Calculating 95% of standardized residuals between -1.96 and 1.96 for outliers for model 2
dataset$fitted <- model_bp$fitted
dataset$residuals <- model_bp$residuals
dataset$standardized.residuals <- rstandard(model_bp)
possible.outliers_model_bp <- subset(dataset, standardized.residuals < -1.96 | standardized.residuals > 1.96)
possible.outliers_model_bp
```

```{r}
# Cooks distance for influential cases for model 2
dataset$cooks <- cooks.distance(model_bp)
plot(sort(dataset$cooks, decreasing=TRUE))
```
```{r}
max(dataset$cooks)
````

